---
title: "Leaflet"
author: "Ashlee, Margarita, Jocelyn, and Rob"
date: "4/12/2022"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE,message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction 
To begin working with Leaflet you will first need to download the package. Use install.packages("leaflet") download this package and load it onto your working script or console.
Leaflet is an open source JavaScript library used to build web mapping applications. It is designed for simplicity, performance and usability. 

# Libraries 
```{r, warning=FALSE, message=FALSE}
library(leaflet)
library(tidyverse)
library(dplyr)
```


# Data 
The data we will be using for our examples is based off of a Tidy Tuesday Dataset on [Volcano Eruptions](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-05-12/readme.md)
```{r, warning=FALSE,message=FALSE}
# Get the Data

volcano <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-12/volcano.csv')
eruptions <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-12/eruptions.csv')
events <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-12/events.csv')
tree_rings <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-12/tree_rings.csv')
sulfur <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-12/sulfur.csv')

```
Wrangling Data 
```{r}
data <- full_join(volcano, eruptions) %>% 
  filter(last_eruption_year >= "1900") %>% 
  select(-major_rock_1, -major_rock_2, -major_rock_3, -major_rock_4, -major_rock_5,
         -minor_rock_1, -minor_rock_2, -minor_rock_3, -minor_rock_4, -minor_rock_5,
         -population_within_5_km, -population_within_10_km, -population_within_30_km, -population_within_100_km,
         -vei) %>% 
  drop_na()
```



# Base Map
Creating a standard basemap 

```{r, warning=FALSE, message=FALSE}

standard.basemap <- leaflet() %>% # leaflet 
 addTiles() %>% 
setView(lat= 37.5, lng=-95.5, zoom=3)

standard.basemap

```

Creating third-party basemaps can be done by searching for different colors and styles of map tiles. 
```{r}

blackandwhite.basemap <- standard.basemap %>% # leaflet 
addProviderTiles(providers$Stamen.Toner)

cartoon.basemap <- standard.basemap %>% # leaflet 
addProviderTiles(providers$CartoDB.Positron)

```

Combining Tile Layers 

```{r}

 basemap <- standard.basemap %>% 
  addProviderTiles(providers$Stamen.Watercolor) %>%
  addProviderTiles(providers$Stamen.TonerLines,
    options = providerTileOptions(opacity = 0.35)) %>% #change opacity
  addProviderTiles(providers$Stamen.TonerLabels)

 basemap

```

Adding Markers

```{r}

longitude <- data$longitude
latitude <- data$latitude
volcanonames <- data$volcano_name

volcanomap <- basemap %>% 
addCircleMarkers(lng = longitude, lat = latitude, popup = volcanonames, radius=3, color="#ffffd4", stroke=FALSE, fillOpacity=1) # add a single point layer

volcanomap

```


